Draft_Application_Stochastic_Volatility

### 2021-07-19 Real Dataset SV.Rmd
update

We start by setting N=50000 and using the Bootstrap particle filter as the benchmark. In table XX we report the RMSE and MAE measures one obtains by comparing the mean of the particles generated at each step by the Bootstrap PF with N=50000 and the mean of the particles generated by other particles filters with N equal to 10, 100, 1000, 5000 or 10000.\footnote{We consider 5 algorithms: a Bootstrap PF, a Guided PF (with the optimal proposal kernel), an Auxiliary PF, a "Basic PF" and a Sequential Importance Sampling alogrithm. By "Basic PF we refer to a Bootstrap PF that features resampling at each step, whereas resampling never happens in the SIS algorithm: one can then consider these 2 algorithms as extreme cases of an adaptive resampling algorithm, namely one that features resampling when ESS is non negative (\textit{i.e.} always) or when ESS<0 (\textit{i.e.} never). In this sense, we use the shorthand "particle filters" when referring to all the 5 algorithms.}

>>>>

We start by running the Bootstrap particle filter with N=50000, setting it as the benchmark. In table XX we report the RMSE and MAE measures one obtains by comparing the mean of the particles generated at each step by the Bootstrap PF with N=50000 and the mean of the particles generated by other particles filters with N equal to 10, 100, 1000, 5000 or 10000.\footnote{We consider 5 algorithms: a Bootstrap PF, a Guided PF (with the optimal proposal kernel), an Auxiliary PF, a "Basic PF" and a Sequential Importance Sampling alogrithm. By "Basic PF we refer to a Bootstrap PF that features resampling at each step, whereas resampling never happens in the SIS algorithm: one can then consider these 2 algorithms as extreme cases of an adaptive resampling algorithm, namely one that features resampling when ESS is non negative (\textit{i.e.} always) or when ESS<0 (\textit{i.e.} never). In this sense, we use the shorthand "particle filters" when referring to all the 5 algorithms.}

```{r include=FALSE}
Errorcomp1<-matrix(NA,ncol=6,nrow=10)
colnames(Errorcomp1)<-c("N","BPF", "GPFOPT", "APF", "BAPF","SIS")
rownames(Errorcomp1)<-c("RMSE","MAE","RMSE","MAE","RMSE","MAE","RMSE","MAE","RMSE","MAE")
Errorcomp1[,1]<-c(10,10,100,100,1000,1000,5000,5000,10000,10000)
Errorcomp1[,2]<-c(0.4007204,0.2688306,0.15268841,0.09413161,0.08482937,0.04579035,0.04275829,0.02021174,0.02931792,0.01369827)
Errorcomp1[,3]<-c(0.4123198,0.2757273,0.15152308,0.09076196,0.06776933,0.03282099,0.06018919,0.02622987,0.03052592,0.01399114)
Errorcomp1[,4]<-c(0.5002458,0.3905036,0.2493751,0.1812517,0.122905,0.08131335,0.06863069,0.03401982,0.04399511,0.02529053)
Errorcomp1[,5]<-c(0.4687276,0.3711974,0.2698023,0.1945553,0.12737504,0.07408736,0.08064348,0.03614264,0.0513611,0.02515361)
Errorcomp1[,6]<-c(0.9168639,0.7275327,1.1055065,0.9179852,0.6250516,0.4971839,0.6200359,0.4475265,0.5118927,0.406363)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, results="asis"}
library(kableExtra)
kbl(Errorcomp1[,], align="c", longtable = T, booktabs = T, caption="RMSE and MAE, Bootstrap PF 50000 particles",digits=5)
```

APF

```{r include=FALSE}
Errorcomp2<-matrix(NA,ncol=6,nrow=10)
colnames(Errorcomp2)<-c("N","BPF", "GPFOPT", "APF", "BAPF", "SIS")
rownames(Errorcomp2)<-c("RMSE","MAE","RMSE","MAE","RMSE","MAE","RMSE","MAE","RMSE","MAE")
Errorcomp2[,1]<-c(10,10,100,100,1000,1000,5000,5000,10000,10000)
Errorcomp2[,2]<-c(0.3780316,0.2581917,0.13740936,0.08692049,0.06682471,0.03859926,0.02801391,0.01508258,0.04292242,0.01831794)
Errorcomp2[,3]<-c(0.3888922,0.2655844,0.13385416,0.08092348,0.05924883,0.03055189,0.05669256,0.02719531,0.04300498,0.02021422)
Errorcomp2[,4]<-c(0.4952102,0.3869999,0.228981,0.1717977,0.10257034,0.07240061,0.04911414,0.02690089,0.0280415,0.01949516)
Errorcomp2[,5]<-c(0.4601913,0.3651146,0.2447648,0.1838898,0.10985867,0.06507043,0.06133971,0.02966254,0.03314042,0.01913677)
Errorcomp2[,6]<-c(0.9103835,0.7211173,1.101557,0.914675,0.6150409,0.4895908,0.6117618,0.4419522,0.5054665,0.4022079)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, results="asis"}
library(kableExtra)
kbl(Errorcomp2[,], align="c", longtable = T, booktabs = T, caption="RMSE and MAE, Auxiliary PF 50000 particles",digits=5)
```

APF vs Bootstrap

```{r include=FALSE}
Errorcomp3<-matrix(NA,ncol=2,nrow=1,byrow=TRUE)
colnames(Errorcomp3)<-c("RMSE", "MAE")
rownames(Errorcomp3)<-("Discrepancy")
Errorcomp3[1,]<-c(0.04407704,0.01667898)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, results="asis"}
library(kableExtra)
kbl(Errorcomp3[,], align="c", longtable = T, booktabs = T, caption="Bootstrap vs Auxiliary PF, 50000 particles",digits=5)
```
